---
title: "Homework 3"
author: "Kireeti Mantrala"
date: "9/26/2022"
output: html_document
---

```{r}
library(datasets)
```

## Time Series dataset for UKgas

* The data shows gas consumption in the UK from 1960-1986
* The data is divided quarterly in millions of therms.

## Plot for the dataset

```{r}
plot(UKgas, main = "Gas consumption in the UK", xlab = "Years", ylab = "Million Therms")

```

* This is the timeseries plot for the dataseries with x-axis showing years and y-axis showing gas consumption in million therms
* The plot is a combination of upward trend with seasonal time series throughout all the quarters.
* The Acf prediction for the above time series plot would be a decreasing trend above the confidence interval along with wavy pattern since the series is both trend and seasonal.

## Autocorrelationfunction for the dataset

```{r}
acf(UKgas)
```

* Most of the values are above the confidence interval. Hence, the historical data will make a significant difference in the forecast.
* As predicted the peaks are in the decreasing pattern showing there is a trend in the time series.
* Also, the pattern showns equally wavy nature, showing this also a seasonal time series.

## HoltWinters Filtering

```{r}
ukhw <- HoltWinters(UKgas)
attributes(ukhw)
plot(ukhw)
```

* The HoltWinters filtering shows that most of the prediction was accurately aligned, except durning the period from 1970-1973 where there is a minor deviation.

```{r}
library(forecast)
ukhwf <- forecast(ukhw)
attributes(ukhwf)
```

## Residual Plot

```{r}
plot(ukhwf$residuals)
```

* The residual plot shows that the residuals are very random and unpredictable.

```{r}
plot(ukhwf$level)
plot(ukhwf$mean)
plot(ukhwf$lower)
plot(ukhwf$upper)
plot(ukhwf$x)
plot(ukhwf$fitted)
hist(ukhwf$residuals)
```

* The histogram looks vaguly distributed with zero as centre but it appears to have slight deviations and not a perfect uniform distribution.

```{r}
Acf(ukhwf$residuals)
```

* Almost all the bars are within the confidence interval except for one value.

```{r}
accuracy(ukhwf)
```

* Generally the Root mean square value should be lower, this shows a very high value showing this is not a correct estimate.
* The HoltWinters filtering shows almost accurate prediction, but the accuracy here shows a highr RMSE value, which is contradicting.